---
title: "Reproducible document"
author: "Mette BI"
format: html
---

## Basics of R

create a new chunk to start coding by using bruge control-shift-p, e.g.:

```{r}
1 + 1
```

## setup code chunk

```{r setup}
#| message: false
#| warning: false
library(tidyverse)
post_meal_data <- read_csv2(here::here("data/post-meal-insulin.csv"))

small_post_meal_data <- select(post_meal_data, id, Age, BMI, Group, auc_pg, auc_ins, glykemi)
```

## Showing the data

```{r}
post_meal_data
```

```{r}
glimpse(post_meal_data)
```

```{r}
ggplot(post_meal_data, aes(x = BMI)) +
  geom_histogram()
```

```{r}
#| fig-cap: "Distribution of BMI"
#| label: fig-mettes-histo
ggplot(post_meal_data, aes(x = BMI)) +
  geom_histogram()
```

@fig-mettes-histo

## Discrete plots

```{r}
#| fig-cap: "Group"
#| label: fig-group-discrete
ggplot(post_meal_data, aes(x = Group)) +
  geom_bar()
```

See @fig-group-discrete above for a cool plot!

```{r}
post_meal_data |>
  ggplot(aes(x = Group, fill = as.character(glykemi))) +
  geom_bar(position = position_dodge())
```
## putting two plots side by side

```{r}
ggplot(post_meal_data, aes(x = BMI)) +
  geom_histogram()
```

```{r}
post_meal_data |>
  ggplot(aes(x = Group, fill = as.character(glykemi))) +
  geom_bar()
```

```{r side-by-side-figs}
#| label: fig-bmi-glycemi
#| fig-cap: "BMI and glycemia, side by side"
#| fig-subcap:
#| - "Distribution of BMI"
#| - "Distribution of glycemia"
#| layout-ncol: 2

ggplot(post_meal_data, aes(x = BMI)) +
  geom_histogram()

post_meal_data |>
  ggplot(aes(x = Group, fill = as.character(glykemi))) +
  geom_bar()
```

```{r side-by-side-figs}
#| label: fig-bmi-glycemia
#| fig-cap: "BMI and glycemia, side by side"
#| fig-subcap:
#| - ""
#| - ""
#| layout-ncol: 2

ggplot(post_meal_data, aes(x = BMI)) +
  geom_histogram()

post_meal_data |>
  ggplot(aes(x = Group, fill = as.character(glykemi))) +
  geom_bar()
```

## Plotting two continuous variables

```{r}
#| fig-cap: "Scatterplot of BMI and the AUC for glucose"
#| label: fig-bmi-auc_pg
ggplot(post_meal_data, aes(x = BMI, y = auc_pg)) +
  geom_point() +
  geom_smooth()
```

## Pushing changes to remote aka github

## Pulling changes locally

## How I accessed Github from R and pushed the project to Github

look at 9 in the pre-course material https://guides.rostools.org/connect-github.html

## Selecting columns from dataset

```{r}
select(post_meal_data, Age)

select(post_meal_data, Age, BMI, Weight)

select(post_meal_data, -Group)

select(post_meal_data, starts_with("PG"))

select(post_meal_data, ends_with("DXA"))

select(post_meal_data, contains("0"))
```

## Renaming specific columns

```{r}
small_post_meal_data |>
  rename(auc_glucose = auc_pg)

small_post_meal_data |>
  rename(auc_glucose = auc_pg, auc_insulin = auc_ins)
```
